\chapter{Debugging}

% === Source file: debug/node-issue.md ===
\section{Node + tsx "\textbackslash\{\}\_\textbackslash\{\}\_name is not a function" crash}


\subsection{Summary}


Running OpenClaw via Node with \texttt{tsx} fails at startup with:

\begin{lstlisting}
[openclaw] Failed to start CLI: TypeError: __name is not a function
    at createSubsystemLogger (.../src/logging/subsystem.ts:203:25)
    at .../src/agents/auth-profiles/constants.ts:25:20
\end{lstlisting}


This began after switching dev scripts from Bun to \texttt{tsx} (commit \texttt{2871657e}, 2026-01-06). The same runtime path worked with Bun.

\subsection{Environment}


\begin{itemize}
  \item Node: v25.x (observed on v25.3.0)
  \item tsx: 4.21.0
  \item OS: macOS (repro also likely on other platforms that run Node 25)
\end{itemize}


\subsection{Repro (Node-only)}


\begin{lstlisting}[language=bash]
# in repo root
node --version
pnpm install
node --import tsx src/entry.ts status
\end{lstlisting}


\subsection{Minimal repro in repo}


\begin{lstlisting}[language=bash]
node --import tsx scripts/repro/tsx-name-repro.ts
\end{lstlisting}


\subsection{Node version check}


\begin{itemize}
  \item Node 25.3.0: fails
  \item Node 22.22.0 (Homebrew \texttt{node@22}): fails
  \item Node 24: not installed here yet; needs verification
\end{itemize}


\subsection{Notes / hypothesis}


\begin{itemize}
  \item \texttt{tsx} uses esbuild to transform TS/ESM. esbuild’s \texttt{keepNames} emits a \texttt{\_\_name} helper and wraps function definitions with \texttt{\_\_name(...)}.
  \item The crash indicates \texttt{\_\_name} exists but is not a function at runtime, which implies the helper is missing or overwritten for this module in the Node 25 loader path.
  \item Similar \texttt{\_\_name} helper issues have been reported in other esbuild consumers when the helper is missing or rewritten.
\end{itemize}


\subsection{Regression history}


\begin{itemize}
  \item \texttt{2871657e} (2026-01-06): scripts changed from Bun to tsx to make Bun optional.
  \item Before that (Bun path), \texttt{openclaw status} and \texttt{gateway:watch} worked.
\end{itemize}


\subsection{Workarounds}


\begin{itemize}
  \item Use Bun for dev scripts (current temporary revert).
  \item Use Node + tsc watch, then run compiled output:
\end{itemize}


\begin{lstlisting}[language=bash]
  pnpm exec tsc --watch --preserveWatchOutput
  node --watch openclaw.mjs status
\end{lstlisting}


\begin{itemize}
  \item Confirmed locally: \texttt{pnpm exec tsc -p tsconfig.json} + \texttt{node openclaw.mjs status} works on Node 25.
  \item Disable esbuild keepNames in the TS loader if possible (prevents \texttt{\_\_name} helper insertion); tsx does not currently expose this.
  \item Test Node LTS (22/24) with \texttt{tsx} to see if the issue is Node 25–specific.
\end{itemize}


\subsection{References}


\begin{itemize}
  \item \href{https://opennext.js.org/cloudflare/howtos/keep\_names}{https://opennext.js.org/cloudflare/howtos/keep\_names}
  \item \href{https://esbuild.github.io/api/\#keep-names}{https://esbuild.github.io/api/\#keep-names}
  \item \href{https://github.com/evanw/esbuild/issues/1031}{https://github.com/evanw/esbuild/issues/1031}
\end{itemize}


\subsection{Next steps}


\begin{itemize}
  \item Repro on Node 22/24 to confirm Node 25 regression.
  \item Test \texttt{tsx} nightly or pin to earlier version if a known regression exists.
  \item If reproduces on Node LTS, file a minimal repro upstream with the \texttt{\_\_name} stack trace.
\end{itemize}



\clearpage
