\chapter{Diagnostics}

% === Source file: diagnostics/flags.md ===
\section{Diagnostics Flags}


Diagnostics flags let you enable targeted debug logs without turning on verbose logging everywhere. Flags are opt-in and have no effect unless a subsystem checks them.

\subsection{How it works}


\begin{itemize}
  \item Flags are strings (case-insensitive).
  \item You can enable flags in config or via an env override.
  \item Wildcards are supported:
  \item \texttt{telegram.*} matches \texttt{telegram.http}
  \item \texttt{*} enables all flags
\end{itemize}


\subsection{Enable via config}


\begin{lstlisting}[language=json]
{
  "diagnostics": {
    "flags": ["telegram.http"]
  }
}
\end{lstlisting}


Multiple flags:

\begin{lstlisting}[language=json]
{
  "diagnostics": {
    "flags": ["telegram.http", "gateway.*"]
  }
}
\end{lstlisting}


Restart the gateway after changing flags.

\subsection{Env override (one-off)}


\begin{lstlisting}[language=bash]
OPENCLAW_DIAGNOSTICS=telegram.http,telegram.payload
\end{lstlisting}


Disable all flags:

\begin{lstlisting}[language=bash]
OPENCLAW_DIAGNOSTICS=0
\end{lstlisting}


\subsection{Where logs go}


Flags emit logs into the standard diagnostics log file. By default:

\begin{lstlisting}
/tmp/openclaw/openclaw-YYYY-MM-DD.log
\end{lstlisting}


If you set \texttt{logging.file}, use that path instead. Logs are JSONL (one JSON object per line). Redaction still applies based on \texttt{logging.redactSensitive}.

\subsection{Extract logs}


Pick the latest log file:

\begin{lstlisting}[language=bash]
ls -t /tmp/openclaw/openclaw-*.log | head -n 1
\end{lstlisting}


Filter for Telegram HTTP diagnostics:

\begin{lstlisting}[language=bash]
rg "telegram http error" /tmp/openclaw/openclaw-*.log
\end{lstlisting}


Or tail while reproducing:

\begin{lstlisting}[language=bash]
tail -f /tmp/openclaw/openclaw-$(date +%F).log | rg "telegram http error"
\end{lstlisting}


For remote gateways, you can also use \texttt{openclaw logs --follow} (see \href{/cli/logs}{/cli/logs}).

\subsection{Notes}


\begin{itemize}
  \item If \texttt{logging.level} is set higher than \texttt{warn}, these logs may be suppressed. Default \texttt{info} is fine.
  \item Flags are safe to leave enabled; they only affect log volume for the specific subsystem.
  \item Use \href{/logging}{/logging} to change log destinations, levels, and redaction.
\end{itemize}



\clearpage
