\chapter{安全}

% === 源文件: security/formal-verification.md ===
\section{形式化验证（安全模型）}


本页跟踪 OpenClaw 的\textbf{形式化安全模型}（目前是 TLA+/TLC；根据需要会添加更多）。

\begin{quote}
注意：一些较旧的链接可能引用了以前的项目名称。
\end{quote}


\textbf{目标（北极星）：} 提供机器检查的论证，证明 OpenClaw 在明确假设下执行其
预期的安全策略（授权、会话隔离、工具门控和
配置错误安全）。

\textbf{目前是什么：} 一个可执行的、攻击者驱动的\textbf{安全回归测试套件}：

\begin{itemize}
  \item 每个声明都有一个在有限状态空间上运行的模型检查。
  \item 许多声明有一个配对的\textbf{负面模型}，为现实的 bug 类别生成反例追踪。
\end{itemize}


\textbf{目前还不是什么：} 证明"OpenClaw 在所有方面都是安全的"或完整 TypeScript 实现是正确的。

\subsection{模型存放位置}


模型维护在一个单独的仓库中：\href{https://github.com/vignesh07/openclaw-formal-models}{vignesh07/openclaw-formal-models}。

\subsection{重要注意事项}


\begin{itemize}
  \item 这些是\textbf{模型}，不是完整的 TypeScript 实现。模型和代码之间可能存在偏差。
  \item 结果受 TLC 探索的状态空间限制；"绿色"并不意味着在建模的假设和边界之外也是安全的。
  \item 一些声明依赖于明确的环境假设（例如，正确的部署、正确的配置输入）。
\end{itemize}


\subsection{复现结果}


目前，结果通过在本地克隆模型仓库并运行 TLC 来复现（见下文）。未来的迭代可能提供：

\begin{itemize}
  \item 带有公开产物（反例追踪、运行日志）的 CI 运行模型
  \item 用于小型、有界检查的托管"运行此模型"工作流
\end{itemize}


开始使用：

\begin{lstlisting}[language=bash]
git clone https://github.com/vignesh07/openclaw-formal-models
cd openclaw-formal-models

# 需要 Java 11+（TLC 在 JVM 上运行）。
# 仓库内置了固定版本的 `tla2tools.jar`（TLA+ 工具）并提供 `bin/tlc` + Make 目标。

make <target>
\end{lstlisting}


\subsubsection{Gateway 网关暴露和开放 Gateway 网关配置错误}


\textbf{声明：} 在没有认证的情况下绑定到 loopback 之外可能使远程入侵成为可能 / 增加暴露；令牌/密码可以阻止未认证的攻击者（根据模型假设）。

\begin{itemize}
  \item 绿色运行：
  \item \texttt{make gateway-exposure-v2}
  \item \texttt{make gateway-exposure-v2-protected}
  \item 红色（预期）：
  \item \texttt{make gateway-exposure-v2-negative}
\end{itemize}


另见：模型仓库中的 \texttt{docs/gateway-exposure-matrix.md}。

\subsubsection{Nodes.run 管道（最高风险能力）}


\textbf{声明：} \texttt{nodes.run} 需要（a）节点命令允许列表加上声明的命令以及（b）配置时的实时批准；批准被令牌化以防止重放（在模型中）。

\begin{itemize}
  \item 绿色运行：
  \item \texttt{make nodes-pipeline}
  \item \texttt{make approvals-token}
  \item 红色（预期）：
  \item \texttt{make nodes-pipeline-negative}
  \item \texttt{make approvals-token-negative}
\end{itemize}


\subsubsection{配对存储（私信门控）}


\textbf{声明：} 配对请求遵守 TTL 和待处理请求上限。

\begin{itemize}
  \item 绿色运行：
  \item \texttt{make pairing}
  \item \texttt{make pairing-cap}
  \item 红色（预期）：
  \item \texttt{make pairing-negative}
  \item \texttt{make pairing-cap-negative}
\end{itemize}


\subsubsection{入站门控（提及 + 控制命令绕过）}


\textbf{声明：} 在需要提及的群组上下文中，未授权的"控制命令"不能绕过提及门控。

\begin{itemize}
  \item 绿色：
  \item \texttt{make ingress-gating}
  \item 红色（预期）：
  \item \texttt{make ingress-gating-negative}
\end{itemize}


\subsubsection{路由/会话键隔离}


\textbf{声明：} 来自不同对等方的私信不会折叠到同一会话中，除非明确链接/配置。

\begin{itemize}
  \item 绿色：
  \item \texttt{make routing-isolation}
  \item 红色（预期）：
  \item \texttt{make routing-isolation-negative}
\end{itemize}


\subsection{v1++：额外的有界模型（并发、重试、追踪正确性）}


这些是后续模型，围绕真实世界的故障模式（非原子更新、重试和消息扇出）提高保真度。

\subsubsection{配对存储并发 / 幂等性}


\textbf{声明：} 配对存储应该在交错情况下也强制执行 \texttt{MaxPending} 和幂等性（即"检查然后写入"必须是原子/加锁的；刷新不应创建重复项）。

这意味着：

\begin{itemize}
  \item 在并发请求下，你不能超过渠道的 \texttt{MaxPending}。
  \item 对同一 \texttt{(channel, sender)} 的重复请求/刷新不应创建重复的活跃待处理行。
\end{itemize}


\begin{itemize}
  \item 绿色运行：
  \item \texttt{make pairing-race}（原子/加锁的上限检查）
  \item \texttt{make pairing-idempotency}
  \item \texttt{make pairing-refresh}
  \item \texttt{make pairing-refresh-race}
  \item 红色（预期）：
  \item \texttt{make pairing-race-negative}（非原子 begin/commit 上限竞争）
  \item \texttt{make pairing-idempotency-negative}
  \item \texttt{make pairing-refresh-negative}
  \item \texttt{make pairing-refresh-race-negative}
\end{itemize}


\subsubsection{入站追踪关联 / 幂等性}


\textbf{声明：} 摄入应在扇出过程中保持追踪关联，并在提供商重试下保持幂等。

这意味着：

\begin{itemize}
  \item 当一个外部事件变成多个内部消息时，每个部分都保持相同的追踪/事件标识。
  \item 重试不会导致重复处理。
  \item 如果提供商事件 ID 缺失，去重回退到安全键（例如，追踪 ID）以避免丢弃不同的事件。
\end{itemize}


\begin{itemize}
  \item 绿色：
  \item \texttt{make ingress-trace}
  \item \texttt{make ingress-trace2}
  \item \texttt{make ingress-idempotency}
  \item \texttt{make ingress-dedupe-fallback}
  \item 红色（预期）：
  \item \texttt{make ingress-trace-negative}
  \item \texttt{make ingress-trace2-negative}
  \item \texttt{make ingress-idempotency-negative}
  \item \texttt{make ingress-dedupe-fallback-negative}
\end{itemize}


\subsubsection{路由 dmScope 优先级 + identityLinks}


\textbf{声明：} 路由必须默认保持私信会话隔离，只有在明确配置时才折叠会话（渠道优先级 + 身份链接）。

这意味着：

\begin{itemize}
  \item 渠道特定的 dmScope 覆盖必须优先于全局默认值。
  \item identityLinks 应该只在明确链接的组内折叠，而不是跨不相关的对等方。
\end{itemize}


\begin{itemize}
  \item 绿色：
  \item \texttt{make routing-precedence}
  \item \texttt{make routing-identitylinks}
  \item 红色（预期）：
  \item \texttt{make routing-precedence-negative}
  \item \texttt{make routing-identitylinks-negative}
\end{itemize}



\clearpage
% === 源文件: security/CONTRIBUTING-THREAT-MODEL.md ===
\section{Contributing to the OpenClaw Threat Model}


Thanks for helping make OpenClaw more secure. This threat model is a living document and we welcome contributions from anyone - you don't need to be a security expert.

\subsection{Ways to Contribute}


\subsubsection{Add a Threat}


Spotted an attack vector or risk we haven't covered? Open an issue on \href{https://github.com/openclaw/trust/issues}{openclaw/trust} and describe it in your own words. You don't need to know any frameworks or fill in every field - just describe the scenario.

\textbf{Helpful to include (but not required):}

\begin{itemize}
  \item The attack scenario and how it could be exploited
  \item Which parts of OpenClaw are affected (CLI, gateway, channels, ClawHub, MCP servers, etc.)
  \item How severe you think it is (low / medium / high / critical)
  \item Any links to related research, CVEs, or real-world examples
\end{itemize}


We'll handle the ATLAS mapping, threat IDs, and risk assessment during review. If you want to include those details, great - but it's not expected.

\begin{quote}
\textbf{This is for adding to the threat model, not reporting live vulnerabilities.} If you've found an exploitable vulnerability, see our \href{https://trust.openclaw.ai}{Trust page} for responsible disclosure instructions.
\end{quote}


\subsubsection{Suggest a Mitigation}


Have an idea for how to address an existing threat? Open an issue or PR referencing the threat. Useful mitigations are specific and actionable - for example, "per-sender rate limiting of 10 messages/minute at the gateway" is better than "implement rate limiting."

\subsubsection{Propose an Attack Chain}


Attack chains show how multiple threats combine into a realistic attack scenario. If you see a dangerous combination, describe the steps and how an attacker would chain them together. A short narrative of how the attack unfolds in practice is more valuable than a formal template.

\subsubsection{Fix or Improve Existing Content}


Typos, clarifications, outdated info, better examples - PRs welcome, no issue needed.

\subsection{What We Use}


\subsubsection{MITRE ATLAS}


This threat model is built on \href{https://atlas.mitre.org/}{MITRE ATLAS} (Adversarial Threat Landscape for AI Systems), a framework designed specifically for AI/ML threats like prompt injection, tool misuse, and agent exploitation. You don't need to know ATLAS to contribute - we map submissions to the framework during review.

\subsubsection{Threat IDs}


Each threat gets an ID like \texttt{T-EXEC-003}. The categories are:


\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Code & Category \\
\hline
RECON & Reconnaissance - information gathering \\
\hline
ACCESS & Initial access - gaining entry \\
\hline
EXEC & Execution - running malicious actions \\
\hline
PERSIST & Persistence - maintaining access \\
\hline
EVADE & Defense evasion - avoiding detection \\
\hline
DISC & Discovery - learning about the environment \\
\hline
EXFIL & Exfiltration - stealing data \\
\hline
IMPACT & Impact - damage or disruption \\
\hline
\end{tabular}
\end{table}



IDs are assigned by maintainers during review. You don't need to pick one.

\subsubsection{Risk Levels}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Level & Meaning \\
\hline
\textbf{Critical} & Full system compromise, or high likelihood + critical impact \\
\hline
\textbf{High} & Significant damage likely, or medium likelihood + critical impact \\
\hline
\textbf{Medium} & Moderate risk, or low likelihood + high impact \\
\hline
\textbf{Low} & Unlikely and limited impact \\
\hline
\end{tabular}
\end{table}



If you're unsure about the risk level, just describe the impact and we'll assess it.

\subsection{Review Process}


\begin{enumerate}
  \item \textbf{Triage} - We review new submissions within 48 hours
  \item \textbf{Assessment} - We verify feasibility, assign ATLAS mapping and threat ID, validate risk level
  \item \textbf{Documentation} - We ensure everything is formatted and complete
  \item \textbf{Merge} - Added to the threat model and visualization
\end{enumerate}


\subsection{Resources}


\begin{itemize}
  \item \href{https://atlas.mitre.org/}{ATLAS Website}
  \item \href{https://atlas.mitre.org/techniques/}{ATLAS Techniques}
  \item \href{https://atlas.mitre.org/studies/}{ATLAS Case Studies}
  \item \href{./THREAT-MODEL-ATLAS.md}{OpenClaw Threat Model}
\end{itemize}


\subsection{Contact}


\begin{itemize}
  \item \textbf{Security vulnerabilities:} See our \href{https://trust.openclaw.ai}{Trust page} for reporting instructions
  \item \textbf{Threat model questions:} Open an issue on \href{https://github.com/openclaw/trust/issues}{openclaw/trust}
  \item \textbf{General chat:} Discord \#security channel
\end{itemize}


\subsection{Recognition}


Contributors to the threat model are recognized in the threat model acknowledgments, release notes, and the OpenClaw security hall of fame for significant contributions.


\clearpage

% === 源文件: security/README.md ===
\section{OpenClaw Security \& Trust}


\textbf{Live:} \href{https://trust.openclaw.ai}{trust.openclaw.ai}

\subsection{Documents}


\begin{itemize}
  \item \href{./THREAT-MODEL-ATLAS.md}{Threat Model} - MITRE ATLAS-based threat model for the OpenClaw ecosystem
  \item \href{./CONTRIBUTING-THREAT-MODEL.md}{Contributing to the Threat Model} - How to add threats, mitigations, and attack chains
\end{itemize}


\subsection{Reporting Vulnerabilities}


See the \href{https://trust.openclaw.ai}{Trust page} for full reporting instructions covering all repos.

\subsection{Contact}


\begin{itemize}
  \item \textbf{Jamieson O'Reilly} (\href{https://twitter.com/theonejvo}{@theonejvo}) - Security \& Trust
  \item Discord: \#security channel
\end{itemize}



\clearpage

% === 源文件: security/THREAT-MODEL-ATLAS.md ===
\section{OpenClaw Threat Model v1.0}


\subsection{MITRE ATLAS Framework}


\textbf{Version:} 1.0-draft
\textbf{Last Updated:} 2026-02-04
\textbf{Methodology:} MITRE ATLAS + Data Flow Diagrams
\textbf{Framework:} \href{https://atlas.mitre.org/}{MITRE ATLAS} (Adversarial Threat Landscape for AI Systems)

\subsubsection{Framework Attribution}


This threat model is built on \href{https://atlas.mitre.org/}{MITRE ATLAS}, the industry-standard framework for documenting adversarial threats to AI/ML systems. ATLAS is maintained by \href{https://www.mitre.org/}{MITRE} in collaboration with the AI security community.

\textbf{Key ATLAS Resources:}

\begin{itemize}
  \item \href{https://atlas.mitre.org/techniques/}{ATLAS Techniques}
  \item \href{https://atlas.mitre.org/tactics/}{ATLAS Tactics}
  \item \href{https://atlas.mitre.org/studies/}{ATLAS Case Studies}
  \item \href{https://github.com/mitre-atlas/atlas-data}{ATLAS GitHub}
  \item \href{https://atlas.mitre.org/resources/contribute}{Contributing to ATLAS}
\end{itemize}


\subsubsection{Contributing to This Threat Model}


This is a living document maintained by the OpenClaw community. See \href{./CONTRIBUTING-THREAT-MODEL.md}{CONTRIBUTING-THREAT-MODEL.md} for guidelines on contributing:

\begin{itemize}
  \item Reporting new threats
  \item Updating existing threats
  \item Proposing attack chains
  \item Suggesting mitigations
\end{itemize}


\hrulefill


\subsection{1. Introduction}


\subsubsection{1.1 Purpose}


This threat model documents adversarial threats to the OpenClaw AI agent platform and ClawHub skill marketplace, using the MITRE ATLAS framework designed specifically for AI/ML systems.

\subsubsection{1.2 Scope}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
Component & Included & Notes \\
\hline
OpenClaw Agent Runtime & Yes & Core agent execution, tool calls, sessions \\
\hline
Gateway & Yes & Authentication, routing, channel integration \\
\hline
Channel Integrations & Yes & WhatsApp, Telegram, Discord, Signal, Slack, etc. \\
\hline
ClawHub Marketplace & Yes & Skill publishing, moderation, distribution \\
\hline
MCP Servers & Yes & External tool providers \\
\hline
User Devices & Partial & Mobile apps, desktop clients \\
\hline
\end{tabular}
\end{table}



\subsubsection{1.3 Out of Scope}


Nothing is explicitly out of scope for this threat model.

\hrulefill


\subsection{2. System Architecture}


\subsubsection{2.1 Trust Boundaries}


\begin{lstlisting}
┌─────────────────────────────────────────────────────────────────┐
│                    UNTRUSTED ZONE                                │
│  ┌─────────────┐  ┌─────────────┐  ┌─────────────┐              │
│  │  WhatsApp   │  │  Telegram   │  │   Discord   │  ...         │
│  └──────┬──────┘  └──────┬──────┘  └──────┬──────┘              │
│         │                │                │                      │
└─────────┼────────────────┼────────────────┼──────────────────────┘
          │                │                │
          ▼                ▼                ▼
┌─────────────────────────────────────────────────────────────────┐
│                 TRUST BOUNDARY 1: Channel Access                 │
│  ┌──────────────────────────────────────────────────────────┐   │
│  │                      GATEWAY                              │   │
│  │  • Device Pairing (30s grace period)                      │   │
│  │  • AllowFrom / AllowList validation                       │   │
│  │  • Token/Password/Tailscale auth                          │   │
│  └──────────────────────────────────────────────────────────┘   │
└─────────────────────────────────────────────────────────────────┘
                              │
                              ▼
┌─────────────────────────────────────────────────────────────────┐
│                 TRUST BOUNDARY 2: Session Isolation              │
│  ┌──────────────────────────────────────────────────────────┐   │
│  │                   AGENT SESSIONS                          │   │
│  │  • Session key = agent:channel:peer                       │   │
│  │  • Tool policies per agent                                │   │
│  │  • Transcript logging                                     │   │
│  └──────────────────────────────────────────────────────────┘   │
└─────────────────────────────────────────────────────────────────┘
                              │
                              ▼
┌─────────────────────────────────────────────────────────────────┐
│                 TRUST BOUNDARY 3: Tool Execution                 │
│  ┌──────────────────────────────────────────────────────────┐   │
│  │                  EXECUTION SANDBOX                        │   │
│  │  • Docker sandbox OR Host (exec-approvals)                │   │
│  │  • Node remote execution                                  │   │
│  │  • SSRF protection (DNS pinning + IP blocking)            │   │
│  └──────────────────────────────────────────────────────────┘   │
└─────────────────────────────────────────────────────────────────┘
                              │
                              ▼
┌─────────────────────────────────────────────────────────────────┐
│                 TRUST BOUNDARY 4: External Content               │
│  ┌──────────────────────────────────────────────────────────┐   │
│  │              FETCHED URLs / EMAILS / WEBHOOKS             │   │
│  │  • External content wrapping (XML tags)                   │   │
│  │  • Security notice injection                              │   │
│  └──────────────────────────────────────────────────────────┘   │
└─────────────────────────────────────────────────────────────────┘
                              │
                              ▼
┌─────────────────────────────────────────────────────────────────┐
│                 TRUST BOUNDARY 5: Supply Chain                   │
│  ┌──────────────────────────────────────────────────────────┐   │
│  │                      CLAWHUB                              │   │
│  │  • Skill publishing (semver, SKILL.md required)           │   │
│  │  • Pattern-based moderation flags                         │   │
│  │  • VirusTotal scanning (coming soon)                      │   │
│  │  • GitHub account age verification                        │   │
│  └──────────────────────────────────────────────────────────┘   │
└─────────────────────────────────────────────────────────────────┘
\end{lstlisting}


\subsubsection{2.2 Data Flows}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|l|l|}
\hline
Flow & Source & Destination & Data & Protection \\
\hline
F1 & Channel & Gateway & User messages & TLS, AllowFrom \\
\hline
F2 & Gateway & Agent & Routed messages & Session isolation \\
\hline
F3 & Agent & Tools & Tool invocations & Policy enforcement \\
\hline
F4 & Agent & External & web\_fetch requests & SSRF blocking \\
\hline
F5 & ClawHub & Agent & Skill code & Moderation, scanning \\
\hline
F6 & Agent & Channel & Responses & Output filtering \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsection{3. Threat Analysis by ATLAS Tactic}


\subsubsection{3.1 Reconnaissance (AML.TA0002)}


\paragraph{T-RECON-001: Agent Endpoint Discovery}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0006 - Active Scanning \\
\hline
\textbf{Description} & Attacker scans for exposed OpenClaw gateway endpoints \\
\hline
\textbf{Attack Vector} & Network scanning, shodan queries, DNS enumeration \\
\hline
\textbf{Affected Components} & Gateway, exposed API endpoints \\
\hline
\textbf{Current Mitigations} & Tailscale auth option, bind to loopback by default \\
\hline
\textbf{Residual Risk} & Medium - Public gateways discoverable \\
\hline
\textbf{Recommendations} & Document secure deployment, add rate limiting on discovery endpoints \\
\hline
\end{tabular}
\end{table}



\paragraph{T-RECON-002: Channel Integration Probing}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0006 - Active Scanning \\
\hline
\textbf{Description} & Attacker probes messaging channels to identify AI-managed accounts \\
\hline
\textbf{Attack Vector} & Sending test messages, observing response patterns \\
\hline
\textbf{Affected Components} & All channel integrations \\
\hline
\textbf{Current Mitigations} & None specific \\
\hline
\textbf{Residual Risk} & Low - Limited value from discovery alone \\
\hline
\textbf{Recommendations} & Consider response timing randomization \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsubsection{3.2 Initial Access (AML.TA0004)}


\paragraph{T-ACCESS-001: Pairing Code Interception}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0040 - AI Model Inference API Access \\
\hline
\textbf{Description} & Attacker intercepts pairing code during 30s grace period \\
\hline
\textbf{Attack Vector} & Shoulder surfing, network sniffing, social engineering \\
\hline
\textbf{Affected Components} & Device pairing system \\
\hline
\textbf{Current Mitigations} & 30s expiry, codes sent via existing channel \\
\hline
\textbf{Residual Risk} & Medium - Grace period exploitable \\
\hline
\textbf{Recommendations} & Reduce grace period, add confirmation step \\
\hline
\end{tabular}
\end{table}



\paragraph{T-ACCESS-002: AllowFrom Spoofing}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0040 - AI Model Inference API Access \\
\hline
\textbf{Description} & Attacker spoofs allowed sender identity in channel \\
\hline
\textbf{Attack Vector} & Depends on channel - phone number spoofing, username impersonation \\
\hline
\textbf{Affected Components} & AllowFrom validation per channel \\
\hline
\textbf{Current Mitigations} & Channel-specific identity verification \\
\hline
\textbf{Residual Risk} & Medium - Some channels vulnerable to spoofing \\
\hline
\textbf{Recommendations} & Document channel-specific risks, add cryptographic verification where possible \\
\hline
\end{tabular}
\end{table}



\paragraph{T-ACCESS-003: Token Theft}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0040 - AI Model Inference API Access \\
\hline
\textbf{Description} & Attacker steals authentication tokens from config files \\
\hline
\textbf{Attack Vector} & Malware, unauthorized device access, config backup exposure \\
\hline
\textbf{Affected Components} & \textasciitilde{}/.openclaw/credentials/, config storage \\
\hline
\textbf{Current Mitigations} & File permissions \\
\hline
\textbf{Residual Risk} & High - Tokens stored in plaintext \\
\hline
\textbf{Recommendations} & Implement token encryption at rest, add token rotation \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsubsection{3.3 Execution (AML.TA0005)}


\paragraph{T-EXEC-001: Direct Prompt Injection}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0051.000 - LLM Prompt Injection: Direct \\
\hline
\textbf{Description} & Attacker sends crafted prompts to manipulate agent behavior \\
\hline
\textbf{Attack Vector} & Channel messages containing adversarial instructions \\
\hline
\textbf{Affected Components} & Agent LLM, all input surfaces \\
\hline
\textbf{Current Mitigations} & Pattern detection, external content wrapping \\
\hline
\textbf{Residual Risk} & Critical - Detection only, no blocking; sophisticated attacks bypass \\
\hline
\textbf{Recommendations} & Implement multi-layer defense, output validation, user confirmation for sensitive actions \\
\hline
\end{tabular}
\end{table}



\paragraph{T-EXEC-002: Indirect Prompt Injection}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0051.001 - LLM Prompt Injection: Indirect \\
\hline
\textbf{Description} & Attacker embeds malicious instructions in fetched content \\
\hline
\textbf{Attack Vector} & Malicious URLs, poisoned emails, compromised webhooks \\
\hline
\textbf{Affected Components} & web\_fetch, email ingestion, external data sources \\
\hline
\textbf{Current Mitigations} & Content wrapping with XML tags and security notice \\
\hline
\textbf{Residual Risk} & High - LLM may ignore wrapper instructions \\
\hline
\textbf{Recommendations} & Implement content sanitization, separate execution contexts \\
\hline
\end{tabular}
\end{table}



\paragraph{T-EXEC-003: Tool Argument Injection}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0051.000 - LLM Prompt Injection: Direct \\
\hline
\textbf{Description} & Attacker manipulates tool arguments through prompt injection \\
\hline
\textbf{Attack Vector} & Crafted prompts that influence tool parameter values \\
\hline
\textbf{Affected Components} & All tool invocations \\
\hline
\textbf{Current Mitigations} & Exec approvals for dangerous commands \\
\hline
\textbf{Residual Risk} & High - Relies on user judgment \\
\hline
\textbf{Recommendations} & Implement argument validation, parameterized tool calls \\
\hline
\end{tabular}
\end{table}



\paragraph{T-EXEC-004: Exec Approval Bypass}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0043 - Craft Adversarial Data \\
\hline
\textbf{Description} & Attacker crafts commands that bypass approval allowlist \\
\hline
\textbf{Attack Vector} & Command obfuscation, alias exploitation, path manipulation \\
\hline
\textbf{Affected Components} & exec-approvals.ts, command allowlist \\
\hline
\textbf{Current Mitigations} & Allowlist + ask mode \\
\hline
\textbf{Residual Risk} & High - No command sanitization \\
\hline
\textbf{Recommendations} & Implement command normalization, expand blocklist \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsubsection{3.4 Persistence (AML.TA0006)}


\paragraph{T-PERSIST-001: Malicious Skill Installation}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0010.001 - Supply Chain Compromise: AI Software \\
\hline
\textbf{Description} & Attacker publishes malicious skill to ClawHub \\
\hline
\textbf{Attack Vector} & Create account, publish skill with hidden malicious code \\
\hline
\textbf{Affected Components} & ClawHub, skill loading, agent execution \\
\hline
\textbf{Current Mitigations} & GitHub account age verification, pattern-based moderation flags \\
\hline
\textbf{Residual Risk} & Critical - No sandboxing, limited review \\
\hline
\textbf{Recommendations} & VirusTotal integration (in progress), skill sandboxing, community review \\
\hline
\end{tabular}
\end{table}



\paragraph{T-PERSIST-002: Skill Update Poisoning}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0010.001 - Supply Chain Compromise: AI Software \\
\hline
\textbf{Description} & Attacker compromises popular skill and pushes malicious update \\
\hline
\textbf{Attack Vector} & Account compromise, social engineering of skill owner \\
\hline
\textbf{Affected Components} & ClawHub versioning, auto-update flows \\
\hline
\textbf{Current Mitigations} & Version fingerprinting \\
\hline
\textbf{Residual Risk} & High - Auto-updates may pull malicious versions \\
\hline
\textbf{Recommendations} & Implement update signing, rollback capability, version pinning \\
\hline
\end{tabular}
\end{table}



\paragraph{T-PERSIST-003: Agent Configuration Tampering}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0010.002 - Supply Chain Compromise: Data \\
\hline
\textbf{Description} & Attacker modifies agent configuration to persist access \\
\hline
\textbf{Attack Vector} & Config file modification, settings injection \\
\hline
\textbf{Affected Components} & Agent config, tool policies \\
\hline
\textbf{Current Mitigations} & File permissions \\
\hline
\textbf{Residual Risk} & Medium - Requires local access \\
\hline
\textbf{Recommendations} & Config integrity verification, audit logging for config changes \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsubsection{3.5 Defense Evasion (AML.TA0007)}


\paragraph{T-EVADE-001: Moderation Pattern Bypass}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0043 - Craft Adversarial Data \\
\hline
\textbf{Description} & Attacker crafts skill content to evade moderation patterns \\
\hline
\textbf{Attack Vector} & Unicode homoglyphs, encoding tricks, dynamic loading \\
\hline
\textbf{Affected Components} & ClawHub moderation.ts \\
\hline
\textbf{Current Mitigations} & Pattern-based FLAG\_RULES \\
\hline
\textbf{Residual Risk} & High - Simple regex easily bypassed \\
\hline
\textbf{Recommendations} & Add behavioral analysis (VirusTotal Code Insight), AST-based detection \\
\hline
\end{tabular}
\end{table}



\paragraph{T-EVADE-002: Content Wrapper Escape}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0043 - Craft Adversarial Data \\
\hline
\textbf{Description} & Attacker crafts content that escapes XML wrapper context \\
\hline
\textbf{Attack Vector} & Tag manipulation, context confusion, instruction override \\
\hline
\textbf{Affected Components} & External content wrapping \\
\hline
\textbf{Current Mitigations} & XML tags + security notice \\
\hline
\textbf{Residual Risk} & Medium - Novel escapes discovered regularly \\
\hline
\textbf{Recommendations} & Multiple wrapper layers, output-side validation \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsubsection{3.6 Discovery (AML.TA0008)}


\paragraph{T-DISC-001: Tool Enumeration}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0040 - AI Model Inference API Access \\
\hline
\textbf{Description} & Attacker enumerates available tools through prompting \\
\hline
\textbf{Attack Vector} & "What tools do you have?" style queries \\
\hline
\textbf{Affected Components} & Agent tool registry \\
\hline
\textbf{Current Mitigations} & None specific \\
\hline
\textbf{Residual Risk} & Low - Tools generally documented \\
\hline
\textbf{Recommendations} & Consider tool visibility controls \\
\hline
\end{tabular}
\end{table}



\paragraph{T-DISC-002: Session Data Extraction}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0040 - AI Model Inference API Access \\
\hline
\textbf{Description} & Attacker extracts sensitive data from session context \\
\hline
\textbf{Attack Vector} & "What did we discuss?" queries, context probing \\
\hline
\textbf{Affected Components} & Session transcripts, context window \\
\hline
\textbf{Current Mitigations} & Session isolation per sender \\
\hline
\textbf{Residual Risk} & Medium - Within-session data accessible \\
\hline
\textbf{Recommendations} & Implement sensitive data redaction in context \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsubsection{3.7 Collection \& Exfiltration (AML.TA0009, AML.TA0010)}


\paragraph{T-EXFIL-001: Data Theft via web\_fetch}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0009 - Collection \\
\hline
\textbf{Description} & Attacker exfiltrates data by instructing agent to send to external URL \\
\hline
\textbf{Attack Vector} & Prompt injection causing agent to POST data to attacker server \\
\hline
\textbf{Affected Components} & web\_fetch tool \\
\hline
\textbf{Current Mitigations} & SSRF blocking for internal networks \\
\hline
\textbf{Residual Risk} & High - External URLs permitted \\
\hline
\textbf{Recommendations} & Implement URL allowlisting, data classification awareness \\
\hline
\end{tabular}
\end{table}



\paragraph{T-EXFIL-002: Unauthorized Message Sending}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0009 - Collection \\
\hline
\textbf{Description} & Attacker causes agent to send messages containing sensitive data \\
\hline
\textbf{Attack Vector} & Prompt injection causing agent to message attacker \\
\hline
\textbf{Affected Components} & Message tool, channel integrations \\
\hline
\textbf{Current Mitigations} & Outbound messaging gating \\
\hline
\textbf{Residual Risk} & Medium - Gating may be bypassed \\
\hline
\textbf{Recommendations} & Require explicit confirmation for new recipients \\
\hline
\end{tabular}
\end{table}



\paragraph{T-EXFIL-003: Credential Harvesting}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0009 - Collection \\
\hline
\textbf{Description} & Malicious skill harvests credentials from agent context \\
\hline
\textbf{Attack Vector} & Skill code reads environment variables, config files \\
\hline
\textbf{Affected Components} & Skill execution environment \\
\hline
\textbf{Current Mitigations} & None specific to skills \\
\hline
\textbf{Residual Risk} & Critical - Skills run with agent privileges \\
\hline
\textbf{Recommendations} & Skill sandboxing, credential isolation \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsubsection{3.8 Impact (AML.TA0011)}


\paragraph{T-IMPACT-001: Unauthorized Command Execution}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0031 - Erode AI Model Integrity \\
\hline
\textbf{Description} & Attacker executes arbitrary commands on user system \\
\hline
\textbf{Attack Vector} & Prompt injection combined with exec approval bypass \\
\hline
\textbf{Affected Components} & Bash tool, command execution \\
\hline
\textbf{Current Mitigations} & Exec approvals, Docker sandbox option \\
\hline
\textbf{Residual Risk} & Critical - Host execution without sandbox \\
\hline
\textbf{Recommendations} & Default to sandbox, improve approval UX \\
\hline
\end{tabular}
\end{table}



\paragraph{T-IMPACT-002: Resource Exhaustion (DoS)}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0031 - Erode AI Model Integrity \\
\hline
\textbf{Description} & Attacker exhausts API credits or compute resources \\
\hline
\textbf{Attack Vector} & Automated message flooding, expensive tool calls \\
\hline
\textbf{Affected Components} & Gateway, agent sessions, API provider \\
\hline
\textbf{Current Mitigations} & None \\
\hline
\textbf{Residual Risk} & High - No rate limiting \\
\hline
\textbf{Recommendations} & Implement per-sender rate limits, cost budgets \\
\hline
\end{tabular}
\end{table}



\paragraph{T-IMPACT-003: Reputation Damage}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Attribute & Value \\
\hline
\textbf{ATLAS ID} & AML.T0031 - Erode AI Model Integrity \\
\hline
\textbf{Description} & Attacker causes agent to send harmful/offensive content \\
\hline
\textbf{Attack Vector} & Prompt injection causing inappropriate responses \\
\hline
\textbf{Affected Components} & Output generation, channel messaging \\
\hline
\textbf{Current Mitigations} & LLM provider content policies \\
\hline
\textbf{Residual Risk} & Medium - Provider filters imperfect \\
\hline
\textbf{Recommendations} & Output filtering layer, user controls \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsection{4. ClawHub Supply Chain Analysis}


\subsubsection{4.1 Current Security Controls}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
Control & Implementation & Effectiveness \\
\hline
GitHub Account Age & \texttt{requireGitHubAccountAge()} & Medium - Raises bar for new attackers \\
\hline
Path Sanitization & \texttt{sanitizePath()} & High - Prevents path traversal \\
\hline
File Type Validation & \texttt{isTextFile()} & Medium - Only text files, but can still be malicious \\
\hline
Size Limits & 50MB total bundle & High - Prevents resource exhaustion \\
\hline
Required SKILL.md & Mandatory readme & Low security value - Informational only \\
\hline
Pattern Moderation & FLAG\_RULES in moderation.ts & Low - Easily bypassed \\
\hline
Moderation Status & \texttt{moderationStatus} field & Medium - Manual review possible \\
\hline
\end{tabular}
\end{table}



\subsubsection{4.2 Moderation Flag Patterns}


Current patterns in \texttt{moderation.ts}:

\begin{lstlisting}[language=javascript]
// Known-bad identifiers
/(keepcold131\/ClawdAuthenticatorTool|ClawdAuthenticatorTool)/i

// Suspicious keywords
/(malware|stealer|phish|phishing|keylogger)/i
/(api[-_ ]?key|token|password|private key|secret)/i
/(wallet|seed phrase|mnemonic|crypto)/i
/(discord\.gg|webhook|hooks\.slack)/i
/(curl[^\n]+\|\s*(sh|bash))/i
/(bit\.ly|tinyurl\.com|t\.co|goo\.gl|is\.gd)/i
\end{lstlisting}


\textbf{Limitations:}

\begin{itemize}
  \item Only checks slug, displayName, summary, frontmatter, metadata, file paths
  \item Does not analyze actual skill code content
  \item Simple regex easily bypassed with obfuscation
  \item No behavioral analysis
\end{itemize}


\subsubsection{4.3 Planned Improvements}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
Improvement & Status & Impact \\
\hline
VirusTotal Integration & In Progress & High - Code Insight behavioral analysis \\
\hline
Community Reporting & Partial (\texttt{skillReports} table exists) & Medium \\
\hline
Audit Logging & Partial (\texttt{auditLogs} table exists) & Medium \\
\hline
Badge System & Implemented & Medium - \texttt{highlighted}, \texttt{official}, \texttt{deprecated}, \texttt{redactionApproved} \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsection{5. Risk Matrix}


\subsubsection{5.1 Likelihood vs Impact}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|l|l|}
\hline
Threat ID & Likelihood & Impact & Risk Level & Priority \\
\hline
T-EXEC-001 & High & Critical & \textbf{Critical} & P0 \\
\hline
T-PERSIST-001 & High & Critical & \textbf{Critical} & P0 \\
\hline
T-EXFIL-003 & Medium & Critical & \textbf{Critical} & P0 \\
\hline
T-IMPACT-001 & Medium & Critical & \textbf{High} & P1 \\
\hline
T-EXEC-002 & High & High & \textbf{High} & P1 \\
\hline
T-EXEC-004 & Medium & High & \textbf{High} & P1 \\
\hline
T-ACCESS-003 & Medium & High & \textbf{High} & P1 \\
\hline
T-EXFIL-001 & Medium & High & \textbf{High} & P1 \\
\hline
T-IMPACT-002 & High & Medium & \textbf{High} & P1 \\
\hline
T-EVADE-001 & High & Medium & \textbf{Medium} & P2 \\
\hline
T-ACCESS-001 & Low & High & \textbf{Medium} & P2 \\
\hline
T-ACCESS-002 & Low & High & \textbf{Medium} & P2 \\
\hline
T-PERSIST-002 & Low & High & \textbf{Medium} & P2 \\
\hline
\end{tabular}
\end{table}



\subsubsection{5.2 Critical Path Attack Chains}


\textbf{Attack Chain 1: Skill-Based Data Theft}

\begin{lstlisting}
T-PERSIST-001 → T-EVADE-001 → T-EXFIL-003
(Publish malicious skill) → (Evade moderation) → (Harvest credentials)
\end{lstlisting}


\textbf{Attack Chain 2: Prompt Injection to RCE}

\begin{lstlisting}
T-EXEC-001 → T-EXEC-004 → T-IMPACT-001
(Inject prompt) → (Bypass exec approval) → (Execute commands)
\end{lstlisting}


\textbf{Attack Chain 3: Indirect Injection via Fetched Content}

\begin{lstlisting}
T-EXEC-002 → T-EXFIL-001 → External exfiltration
(Poison URL content) → (Agent fetches & follows instructions) → (Data sent to attacker)
\end{lstlisting}


\hrulefill


\subsection{6. Recommendations Summary}


\subsubsection{6.1 Immediate (P0)}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
ID & Recommendation & Addresses \\
\hline
R-001 & Complete VirusTotal integration & T-PERSIST-001, T-EVADE-001 \\
\hline
R-002 & Implement skill sandboxing & T-PERSIST-001, T-EXFIL-003 \\
\hline
R-003 & Add output validation for sensitive actions & T-EXEC-001, T-EXEC-002 \\
\hline
\end{tabular}
\end{table}



\subsubsection{6.2 Short-term (P1)}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
ID & Recommendation & Addresses \\
\hline
R-004 & Implement rate limiting & T-IMPACT-002 \\
\hline
R-005 & Add token encryption at rest & T-ACCESS-003 \\
\hline
R-006 & Improve exec approval UX and validation & T-EXEC-004 \\
\hline
R-007 & Implement URL allowlisting for web\_fetch & T-EXFIL-001 \\
\hline
\end{tabular}
\end{table}



\subsubsection{6.3 Medium-term (P2)}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
ID & Recommendation & Addresses \\
\hline
R-008 & Add cryptographic channel verification where possible & T-ACCESS-002 \\
\hline
R-009 & Implement config integrity verification & T-PERSIST-003 \\
\hline
R-010 & Add update signing and version pinning & T-PERSIST-002 \\
\hline
\end{tabular}
\end{table}



\hrulefill


\subsection{7. Appendices}


\subsubsection{7.1 ATLAS Technique Mapping}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
ATLAS ID & Technique Name & OpenClaw Threats \\
\hline
AML.T0006 & Active Scanning & T-RECON-001, T-RECON-002 \\
\hline
AML.T0009 & Collection & T-EXFIL-001, T-EXFIL-002, T-EXFIL-003 \\
\hline
AML.T0010.001 & Supply Chain: AI Software & T-PERSIST-001, T-PERSIST-002 \\
\hline
AML.T0010.002 & Supply Chain: Data & T-PERSIST-003 \\
\hline
AML.T0031 & Erode AI Model Integrity & T-IMPACT-001, T-IMPACT-002, T-IMPACT-003 \\
\hline
AML.T0040 & AI Model Inference API Access & T-ACCESS-001, T-ACCESS-002, T-ACCESS-003, T-DISC-001, T-DISC-002 \\
\hline
AML.T0043 & Craft Adversarial Data & T-EXEC-004, T-EVADE-001, T-EVADE-002 \\
\hline
AML.T0051.000 & LLM Prompt Injection: Direct & T-EXEC-001, T-EXEC-003 \\
\hline
AML.T0051.001 & LLM Prompt Injection: Indirect & T-EXEC-002 \\
\hline
\end{tabular}
\end{table}



\subsubsection{7.2 Key Security Files}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|l|}
\hline
Path & Purpose & Risk Level \\
\hline
\texttt{src/infra/exec-approvals.ts} & Command approval logic & \textbf{Critical} \\
\hline
\texttt{src/gateway/auth.ts} & Gateway authentication & \textbf{Critical} \\
\hline
\texttt{src/web/inbound/access-control.ts} & Channel access control & \textbf{Critical} \\
\hline
\texttt{src/infra/net/ssrf.ts} & SSRF protection & \textbf{Critical} \\
\hline
\texttt{src/security/external-content.ts} & Prompt injection mitigation & \textbf{Critical} \\
\hline
\texttt{src/agents/sandbox/tool-policy.ts} & Tool policy enforcement & \textbf{Critical} \\
\hline
\texttt{convex/lib/moderation.ts} & ClawHub moderation & \textbf{High} \\
\hline
\texttt{convex/lib/skillPublish.ts} & Skill publishing flow & \textbf{High} \\
\hline
\texttt{src/routing/resolve-route.ts} & Session isolation & \textbf{Medium} \\
\hline
\end{tabular}
\end{table}



\subsubsection{7.3 Glossary}



\begin{table}[h]
\centering
\small
\begin{tabular}{|l|l|}
\hline
Term & Definition \\
\hline
\textbf{ATLAS} & MITRE's Adversarial Threat Landscape for AI Systems \\
\hline
\textbf{ClawHub} & OpenClaw's skill marketplace \\
\hline
\textbf{Gateway} & OpenClaw's message routing and authentication layer \\
\hline
\textbf{MCP} & Model Context Protocol - tool provider interface \\
\hline
\textbf{Prompt Injection} & Attack where malicious instructions are embedded in input \\
\hline
\textbf{Skill} & Downloadable extension for OpenClaw agents \\
\hline
\textbf{SSRF} & Server-Side Request Forgery \\
\hline
\end{tabular}
\end{table}



\hrulefill


\_This threat model is a living document. Report security issues to security@openclaw.ai\_


\clearpage

